---
title: "SQL vs. dbplyr"
output:
  html_document: default
  html_notebook: default
---

## Getting data

Download data `database.sqlite.zip` from  <https://www.kaggle.com/hugomathien/soccer> into directory 
`football_data` and uznip it.

## Access to data base

```{r}
DATA_DIR <- file.path("..", "football_data")
library(RSQLite)
sql_con <- dbConnect(
  SQLite(), 
  dbname=file.path(DATA_DIR, "database.sqlite")
)
dbListTables(sql_con)
```

Example of query:
```{r}
dbGetQuery(sql_con, "
SELECT *
FROM Player
LIMIT 6
")
```

sql chunk:
```{sql, connection=sql_con}
SELECT *
FROM Player_Attributes
LIMIT 6
```

## dplyr 

We will show how to work with dplyr with remote storage. This will connect data base with dplyr.
```{r, results=FALSE, echo=FALSE}
library('dplyr')
player_db <- tbl(sql_con, "Player")
player_attributes_db <- tbl(sql_con, "Player_Attributes")
match_db <- tbl(sql_con, "Match")
```

### Pipes

First let us introduce pipes.
```{r}
print("Hello world!")
"Hello world!" %>%
  print
```

```{r}
2 + 3 %>%
  print

3 %>%
  {2 + .} %>%
  print
```

## Show query

```{r}
player_db %>%
  mutate(year=as.numeric(strftime('%Y', birthday))) %>%
  group_by(year) %>%
  summarise(height=mean(height, na.rm=TRUE)) %>%
  show_query
```


## SELECT ~ select

```{r}
"SELECT player_name
,   birthday
FROM Player
LIMIT 10" %>%
  dbGetQuery(sql_con, .)
```

```{r}
player_db %>%
  select(player_name, birthday) %>%
  head(10)
```

## WHERE ~ filter

```{sql where_sql, connection=sql_con}
SELECT *
FROM Player
WHERE birthday >= '1998'
LIMIT 5
```

```{r}
player_db %>%
  filter(birthday >= '1998') %>%
  head(5)
```

## ORDER ~ arrange

```{r}
player_db %>%
  arrange(birthday) %>%
  head(5)
```

```{r}
"SELECT *
FROM Player
ORDER BY birthday 
LIMIT 5" %>%
  dbGetQuery(sql_con, .)
```

```{r}
player_db %>%
  arrange(desc(birthday)) %>%
  head(5)
```

```{r}
"SELECT *
FROM Player
ORDER BY birthday DESC
LIMIT 5" %>%
  dbGetQuery(sql_con, .)
```
## AS ~ mutate

```{r}
player_db %>%
  mutate(weight_kg = weight*0.45359237) %>%
  head(3) %>%
  data.frame
```

```{r}
"SELECT *
,   weight*0.45359237 AS weight_kg
FROM Player
LIMIT 3" %>%
  dbGetQuery(sql_con, .)
```

## aggregators ~ summarise

```{r}
player_db %>%
  summarise(
    mean_weight=mean(weight, na.rm=TRUE),
    sd_weight=sd(weight, na.rm=TRUE),
    sum_weight=sum(weight, na.rm=TRUE),
    n_samples=n())
```

```{r}
"SELECT AVG(weight) AS mean_weight
,   STDEV(weight) AS sd_weight
,   SUM(weight) AS sum_weight
,   COUNT(1) AS n_samples
FROM Player
" %>%
  dbGetQuery(sql_con, .)
```

## GROUP BY ~ group_by

Here we use SQLite date functions
<https://www.sqlite.org/lang_datefunc.html>

```{r}
library('ggplot2')
"SELECT CAST(STRFTIME('%Y', birthday) AS INT) AS year
,   AVG(height) AS height
FROM Player
GROUP BY CAST(SUBSTR(birthday, 1, 4) AS INT) " %>%
  dbGetQuery(sql_con, .) %>%
  ggplot(., aes(year, height)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```

```{r}
player_db %>%
  mutate(year=as.numeric(strftime('%Y', birthday))) %>%
  group_by(year) %>%
  summarise(height=mean(height, na.rm=TRUE)) %>%
  data.frame %>%
  ggplot(aes(year, height)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```



## Joins

```{sql connection=sql_con}
SELECT Player.player_name
, pa.max_rating 
FROM Player
LEFT JOIN (
  SELECT player_api_id
  ,  MAX(overall_rating) AS max_rating
  FROM Player_Attributes
  GROUP BY player_api_id) pa
ON Player.player_api_id = pa.player_api_id
ORDER BY pa.max_rating DESC
LIMIT 10
```
...

```{r}
player_db <- tbl(sql_con, "Player")
player_attributes_db <- tbl(sql_con, "Player_Attributes")

player_db %>%
  left_join(
    {player_attributes_db %>% 
      group_by(player_api_id) %>%
      summarise(max_rating=max(overall_rating, na.rm=TRUE))}, 
    by='player_api_id') %>%
  arrange(-max_rating) %>%
  select(player_name, max_rating) %>%
  head(10)
```

## UNION All ~ union_all

```{sql, connection=sql_con}
SELECT team_api_id
,   SUM(goal) AS goals
FROM (
    SELECT home_team_api_id AS team_api_id
    ,   home_team_goal AS goal
    FROM Match
    WHERE country_id = 1
    AND league_id = 1    
    AND season = '2008/2009'
    UNION ALL 
    SELECT away_team_api_id AS team_api_id
    ,   away_team_goal AS goal
    FROM Match
    WHERE country_id = 1
    AND league_id = 1    
    AND season = '2008/2009')
GROUP BY 1
ORDER BY 2 DESC
```

```{r}
union_all(
{match_db %>%
  filter(country_id == 1, league_id == 1, season == '2008/2009') %>%
  select(team_api_id = home_team_api_id, team_goal = home_team_goal)},
{match_db %>%
  filter(country_id == 1, league_id == 1, season == '2008/2009') %>%
  select(team_api_id = away_team_api_id, team_goal = away_team_goal)}) %>%
  group_by(team_api_id) %>%
  summarise(goals=sum(team_goal, na.rm=TRUE)) %>%
  arrange(desc(goals))
```


=======

Exercise

* Translate from SQL to R (with dplyr) or from R to SQL
```{r}
"SELECT COUNT(1) AS n
FROM Match" %>%
  dbGetQuery(sql_con, .)
```


```{r}
"SELECT country_id
,   COUNT(1) AS n
FROM Match
GROUP BY country_id
ORDER BY n DESC" %>%
  dbGetQuery(sql_con, .)
```

```{r}
match_db <- tbl_df(dbGetQuery(sql_con,"SELECT * FROM Match"))
country_db <- tbl_df(dbGetQuery(sql_con,"SELECT * FROM Country"))
teams <- tbl_df(dbGetQuery(sql_con,"SELECT * FROM Team"))
```


```{r}
match_db %>%
  left_join(country_db, by=c("country_id" = "id")) %>%
  select(name, country_id, home_team_api_id) %>%
  group_by(name) %>%
  summarise(teams = n_distinct(home_team_api_id)) %>%
  arrange(desc(teams))
```

* List 5 players and their highest rating that had the lowest potential ever.

* Which team has the higest average of goles when playing home in season 2008/2009 in Poland?

* Which team and from which country and which season scored the higest number of goals?

## Whisker

<https://github.com/edwindj/whisker>

```{r}
library(whisker)
template <- "
Hello,
My name is {{{name}}} and I like {{{thing_i_like}}}.
"

data <- list( 
  name="Bartek", 
  thing_i_like="tenis")

text <- whisker.render(template, data)
cat(text)
```

```{r}
template <- "
SELECT *
FROM Player
WHERE player_name = '{{{player_name}}}'
"

data <- list( 
  player_name="Cristiano Ronaldo")

query<- whisker.render(template, data)

dbGetQuery(sql_con, query)

 
```

## Close connection

```{r}
dbDisconnect(sql_con)
```

https://db.rstudio.com/dplyr/
